---
title: "HW9_Survival_Analysis"
author: "Group1"
date: '2021 11 22'
output: 
  html_document: 
    theme: paper
    toc: yes
    toc_float: yes
    df_print: paged
---

```{r include=FALSE}
#setwd("C:/Users/qkrwl/Desktop/2021-2/이통2/HW/hw9/")
library(readxl)
library(tidyverse)
library(gridExtra)
library(GGally)
library(survival)
library(survminer)
library(gtsummary)
library(MASS)
library(fastDummies)
library(ggfortify)
```

```{r}
data <- read_xls("Framingham.xls")
data$sex <- as.factor(data$sex)
data$month <- as.factor(data$month)
head(data)
na.ind <- which(!complete.cases(data))
data <- data[,-1]
```

# PART I. Cox

```{r}
data2 <- data[-na.ind,]
```

```{r}
# ggpairs(data2)

# par(mfrow=c(1,2))
# for(i in 1:5){
#   hist(data2[,i], xlab = "", main=colnames(data2)[i])
#   hist(log(data2[,i]), xlab = "", main=paste0("log(",colnames(data2)[i],")"))
# }

data.t <- data2

# log 변환
data.t$sbp <- log(data.t$sbp)
data.t$bmi <- log(data.t$bmi)
data.t$dbp <- log(data.t$dbp)
data.t$scl <- log(data.t$scl)

# age 그룹화
data.t$age <- as.factor(floor(data.t$age/10)*10)


# centering
col.mean.t <- colMeans(data.t[,c(1:3,5)])
col.mean.t
data.t[,c(1:3,5)] <- scale(data.t[,c(1:3,5)], center = TRUE, scale = FALSE)

half <- list()
col <- c(1:3,5)
for(i in 1:4){
  half[[i]] <- gghalfnorm::gghalfnorm(data.t[[col[i]]], nlab = 3, repel = T)
}
grid.arrange(grobs=half, ncol=2)

out <- c(3360, 956, 1225, 1906, 3329, 3435, 612) 
```

## (a)
```{r message=FALSE, warning=FALSE}
data.t <- data2[-out,]

# log 변환
data.t$sbp <- log(data.t$sbp)
data.t$bmi <- log(data.t$bmi)
data.t$dbp <- log(data.t$dbp)
data.t$scl <- log(data.t$scl)

# age 그룹화
data.t$age <- as.factor(floor(data.t$age/10)*10)


# centering
col.mean.t <- colMeans(data.t[,c(1:3,5)])
col.mean.t
data.t[,c(1:3,5)] <- scale(data.t[,c(1:3,5)], center = TRUE, scale = FALSE)

cox.t <- coxph(Surv(followup, chdfate)~., data = data.t)
cox.t.step <-stepAIC(cox.t,direction = "both", trace = 0)

# intercation
cox.t2 <- coxph(Surv(followup, chdfate)~.*., data = data.t)
cox.t2.step <-stepAIC(cox.t2,direction = "both", trace = 0)

data.frame('simple'=extractAIC(cox.t), 'simplt_step'=extractAIC(cox.t.step), 
           'interaction'=extractAIC(cox.t2), 'interaction_step'=extractAIC(cox.t2.step),
           row.names=c('var_num', 'AIC'))

ggforest(cox.t2.step, data.t)
ggsurvplot(survfit(cox.t2.step), data=data.t, conf.int = FALSE)
cox.t2.step %>% tbl_regression() %>% plot()
```

coxph(formula = Surv(followup, chdfate) ~ sbp + dbp + scl + age + 
    bmi + sex + sbp:dbp + sbp:scl + sbp:age + dbp:scl + dbp:age + 
    dbp:bmi + scl:age, data = data.t)

```{r eval=FALSE, include=FALSE}
# log 변환 x
# centering
col.mean <- colMeans(data2[-na.ind,1:5])
col.mean
data.s <- data2[-na.ind,]
data.s[,1:5] <- scale(data.s[,1:5], center = TRUE, scale = FALSE)


cox.s <- coxph(Surv(followup, chdfate)~., data = data.s)
extractAIC(cox.s)

cox.s.step <-step(cox.s,direction = "both", trace = 0)
extractAIC(cox.s.step)
# ggforest(cox.s.step, data = data.s)
# ggsurvplot(survfit(cox.s.step), data=data.s, conf.int = FALSE)

# intercation
cox.s2 <- coxph(Surv(followup, chdfate)~.*., data = data.s)
extractAIC(cox.s2)

cox.s2.step <-step(cox.s2,direction = "both", trace = 0)
extractAIC(cox.s2.step)

ggforest(cox.s2.step, data = data.s)
ggsurvplot(survfit(cox.s2.step), data=data.s)
summary(cox.s2.step)
```

## (b)
```{r message=FALSE, warning=FALSE}
par(mfrow=c(1,2))
ggsurvplot(survfit(cox.t2.step), data=data.t, conf.int = FALSE)
ggsurvplot(survfit(cox.t2.step), fun="cumhaz", data=data.t, conf.int = FALSE, palette = "skyblue3")
```

```{r message=FALSE, warning=FALSE}
S <- basehaz(cox.t2.step)

par(mfrow=c(1,2))
with(S, plot(time, log(hazard), main="Gompertz"))
with(S, plot(log(time), log(hazard), main="Weibull"))
```

## (c)
```{r message=FALSE, warning=FALSE}
cox.t2.step %>% tbl_regression(exponentiate=T) %>% plot()
autoplot(aareg(cox.t2.step$formula, data.t), ncol=4, scales="free")
```

# PART II. ALT
## a)
```{r warning = FALSE, message = FALSE}
#weibull
alt.w <- survreg(Surv(followup,chdfate)~., data = data.t, dist = 'weibull')
alt.w.i <- survreg(Surv(followup,chdfate)~.*., data = data.t, dist = 'weibull')

alt.w.step <-step(alt.w,trace=F)
alt.w.i.step <-step(alt.w.i,trace=F)

#loglogistic
alt.l <- survreg(Surv(followup,chdfate )~., data = data.t, dist = 'loglogistic')
alt.l.i <- survreg(Surv(followup,chdfate)~.*., data = data.t, dist = 'loglogistic')

alt.l.step<-step(alt.l,trace=F)
alt.l.i.step<-step(alt.l.i,trace=F)

#lognormal
alt.ln <- survreg(Surv(followup,chdfate )~., data = data.t, dist = 'lognormal')
alt.ln.i <- survreg(Surv(followup,chdfate)~.*., data = data.t, dist = 'lognormal')

alt.ln.step<-step(alt.ln,trace=F)
alt.ln.i.step<-step(alt.ln.i,trace=F)

#logistic
alt.log <- survreg(Surv(followup,chdfate )~., data = data.t, dist = 'logistic')
alt.log.i <- survreg(Surv(followup,chdfate )~.*., data = data.t, dist = 'logistic')

alt.log.step<-step(alt.log,trace=F)
alt.log.i.step<-step(alt.log.i,trace=F)


aic.w<-rbind(simple=extractAIC(alt.w),interction=extractAIC(alt.w.i),
             simple.step=extractAIC(alt.w.step),
             interaction.step=extractAIC(alt.w.i.step))

aic.l<-rbind(simple=extractAIC(alt.l),interction=extractAIC(alt.l.i),
             simple.step=extractAIC(alt.l.step),
             interaction.step=extractAIC(alt.l.i.step))

aic.ln<-rbind(simple=extractAIC(alt.ln),interction=extractAIC(alt.ln.i),
              simple.step=extractAIC(alt.ln.step),
              interaction.step=extractAIC(alt.ln.i.step))

aic.log<-rbind(simple=extractAIC(alt.log),interction=extractAIC(alt.log.i),
               simple.step=extractAIC(alt.log.step),
               interaction.step=extractAIC(alt.log.i.step))

aic<-cbind(aic.w,aic.l,aic.ln,aic.log)
colnames(aic) <- c('no', 'weibull', 'no', 'loglogistic',
                   'no', 'lognormal', 'no', 'logistic')
aic <- data.frame(aic)
aic 

# Weibull & interaction term 
# best : alt.w.i.step

alt<-alt.w.i.step
summary(alt)
extractAIC(alt)

sigma <- alt$scale
mu <- predict(alt,type = "lp")

alt.data <- data.frame(u = (log(data.t$followup)-mu)/sigma) %>% mutate(r = rank(u), p = r/(length(data.t$followup)+1), weibull =  log(-log(1-p))) 
ggplot(alt.data, aes(weibull, u)) + geom_point() + geom_smooth(method='lm', se = F) + ggtitle('Weibull Q-Q plot')
```

Weibull을 최적 모형으로 선택함. 

survreg(formula = Surv(followup, chdfate) ~ age + sex + log_scl + 
    log_sbp + log_bmi + log_dbp + age:sex + age:log_sbp + age:log_dbp + 
    log_scl:log_sbp + log_scl:log_dbp + log_sbp:log_dbp + log_bmi:log_dbp, 
    data = df, dist = "weibull")

## b)
```{r warning = FALSE, message = FALSE}
alt.w.i.step %>% tbl_regression() %>% plot()

alt_form <- survreg(formula = Surv(followup, chdfate) ~ sbp + dbp + scl + 
    age + bmi + sex + sbp:dbp + sbp:scl + sbp:age + dbp:scl + 
    dbp:age + dbp:bmi + scl:age, data = data.t, dist = "weibull")

autoplot(aareg(alt_form, data.t), ncol=4, scales="free")
```

# PART3
## (a) & (b)
```{r message=FALSE, warning=FALSE}
# cox
summary(survfit(cox.t2.step), time=3650)

lm <- lm(log(hazard)~log(time), data = S)
alpha <- summary(lm)$coefficients[1]
beta <- summary(lm)$coefficients[2]

mu.cox <- predict(cox.t2.step, data.t, type = 'risk')
p.cox <- 1-exp(-exp((alpha+beta*(log(3650))))*mu.cox)
cox_hist <- hist(p.cox, breaks = 20, xlab="Cox PHM", main="Histogram of Probability (10years)")

# alt
u <- (log(3650)-mu)/sigma
prob_alt <- (1-exp(-exp(u)))
alt_hist <- hist(prob_alt, breaks = 20, xlab="ALT", main="Histogram of Probability (10years)")

# plot
plot(p.cox, prob_alt, main='Probability of Cox vs ALT', xlab='COX', ylab='ALT')

plot(density(prob_alt), col='skyblue', lwd=2, main='Density Probability of Cox vs ALT')
lines(density(p.cox), col='lightsalmon', lwd=2)
legend("topright", legend = c("ALT", "COX"), col=c("skyblue","lightsalmon"), lwd=2, inset=0.05)
```

## (c)
```{r message=FALSE, warning=FALSE}
data.frame('Cox' = extractAIC(cox.t2.step), 'ALT' = extractAIC(alt.w.i.step),
           row.names = c('var.num', 'AIC'))
```

# PART4
### cox
```{r message=FALSE, warning=FALSE}
data.new <- data2[-out,] %>% mutate(t.star = age*365.25+followup) %>% dplyr::select(-followup, -age)
data.new[,1:4] <- scale(data.new[,1:4], center = TRUE, scale = FALSE)

cox.new <- coxph(Surv(t.star, chdfate)~., data = data.new)
cox.new.step <-step(cox.new,direction = "both", trace = 0)

# intercation
cox.new2 <- coxph(Surv(t.star, chdfate)~.*., data = data.new)
cox.new2.step <-step(cox.new2,direction = "both", trace = 0)

cox.new2.step %>% tbl_regression(exponentiate=T) %>% plot()
ggsurvplot(survfit(cox.new2.step), data=data.new)
ggsurvplot(survfit(cox.new2.step), data=data.new, fun = "cumhaz", palette="skyblue3")
#summary(cox.new2.step)

ggsurvplot(survfit(cox.new2.step), data=data.new)

data.frame('simple'=extractAIC(cox.new), 'simple_step'=extractAIC(cox.new.step), 
           'interaction' = extractAIC(cox.new2), 'interaction_step' = extractAIC(cox.new2.step),
           row.names=c('var_num', 'AIC'))

# H(t), S(t)
S2 <- data.frame(time = survfit(cox.new2.step)$time, surv = survfit(cox.new2.step)$surv, hazard = -log(survfit(cox.new2.step)$surv))

par(mfrow=c(1,2))
with(S2, plot(time, log(hazard), main="Gompertz"))
with(S2, plot(log(time), log(hazard), main="Weibull"))
```

```{r}
# probability 예측
lm2 <- lm(log(hazard)~log(time), data = S2[-(1:7),])
alpha2 <- summary(lm2)$coefficients[1]
beta2 <- summary(lm2)$coefficients[2]

mu.cox2 <- predict(cox.new2.step, data.new, type = 'risk')
t <- data2$age[-out]*365.25+3650
p.cox2 <- 1-exp(-exp((alpha2+beta2*log(t)))*mu.cox2)

# plot
hist(p.cox2, breaks = 30,  xlab="age x 365.25 + followup", main="Histogram of COX Probability (10years)")

plot(density(p.cox2), col="skyblue", lwd=2, main="Probability Comparing of COX", xlab="")
lines(density(p.cox), col="lightsalmon", lwd=2)
legend("topright", legend = c("age x 365.25 + followup", "followup"), col=c("skyblue","lightsalmon"), lwd=2, inset=0.05)

plot(p.cox, p.cox2, xlab="followup", ylab="age x 365.25 + followup", main="Probability Comparing of COX")
abline(0,1, col="blue", lty=2)
```

### alt
```{r warning = FALSE, message = FALSE}
#weibull
alt.w <- survreg(Surv(t.star,chdfate)~., data = data.new, dist = 'weibull')
alt.w.i <- survreg(Surv(t.star,chdfate)~.*., data = data.new, dist = 'weibull')

alt.w.step<-step(alt.w,trace=F)
alt.w.i.step<-step(alt.w.i,trace=F)

#loglogistic
alt.l <- survreg(Surv(t.star,chdfate )~., data = data.new, dist = 'loglogistic')
alt.l.i <- survreg(Surv(t.star,chdfate)~.*., data = data.new, dist = 'loglogistic')

alt.l.step<-step(alt.l,trace=F)
alt.l.i.step<-step(alt.l.i,trace=F)

#lognormal
alt.ln <- survreg(Surv(t.star,chdfate )~., data = data.new, dist = 'lognormal')
alt.ln.i <- survreg(Surv(t.star,chdfate)~.*., data = data.new, dist = 'lognormal')

alt.ln.step<-step(alt.ln,trace=F)
alt.ln.i.step<-step(alt.ln.i,trace=F)


#logistic
alt.log <- survreg(Surv(t.star,chdfate )~., data = data.new, dist = 'logistic')
alt.log.i <- survreg(Surv(t.star,chdfate )~.*., data = data.new, dist = 'logistic')

alt.log.step<-step(alt.log,trace=F)
alt.log.i.step<-step(alt.log.i,trace=F)


aic.w<-rbind(simple=AIC(alt.w),interction=AIC(alt.w.i),
             simple.step=AIC(alt.w.step),interaction.step=AIC(alt.w.i.step))

aic.l<-rbind(simple=AIC(alt.l),interction=AIC(alt.l.i),
             simple.step=AIC(alt.l.step),interaction.step=AIC(alt.l.i.step))

aic.ln<-rbind(simple=AIC(alt.ln),interction=AIC(alt.ln.i),
              simple.step=AIC(alt.ln.step),interaction.step=AIC(alt.ln.i.step))

aic.log<-rbind(simple=AIC(alt.log),interction=AIC(alt.log.i),
               simple.step=AIC(alt.log.step),interaction.step=AIC(alt.log.i.step))

aic<-cbind(aic.w,aic.l,aic.ln,aic.log)
colnames(aic) <- c('weibull','loglogistic','lognormal','logistic')
aic <- data.frame(aic)
aic

# lognormal  & interaction term
# best : alt.ln.i.step

alt2<-alt.ln.i.step

sigma2 <- alt2$scale
mu2 <- predict(alt2,type = "lp")

alt.data2 <- data.frame(u =(log(data.new$t.star)-mu2)/sigma2) %>%
  mutate(r = rank(u), p = r/(length(data.new$t.star)+1), lognormal = qnorm(p))
ggplot(alt.data2, aes(lognormal, u)) + geom_point() + geom_smooth(method='lm', se = F) + ggtitle('lognormal Q-Q plot')

u2 <- (log(t)-mu2)/sigma2
prob_alt2 <- pnorm(u2)

# plot
hist(prob_alt2, breaks = 30,  xlab="age x 365.25 + followup", main="Histogram of ALT Probability (10years)")

plot(density(prob_alt2), col="skyblue", lwd=2, main="Probability Comparing of ALT", xlab="")
lines(density(prob_alt), col="lightsalmon", lwd=2)
legend("topright", legend = c("age x 365.25 + followup", "followup"), col=c("skyblue","lightsalmon"), lwd=2, inset=0.05)

plot(prob_alt,prob_alt2, xlab="followup", ylab="age x 365.25 + followup", main="Probability Comparing of ALT")
abline(0,1, col="blue", lty=2)
```

